The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-12-09 Manuscript_draft/Draft.Rmd
Introduction

Forest management have the potential to affect respiration rates by modifying the microclimate. Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it by reducing the transpiration of water by trees (Zhang et al 2018). The degree of the response of soil will be dependent on the severity of the forest intervention. Thus, more severe forestry intervention such as clear-cuts may increase soil moisture, enhance heat and frost damage (Kovaks et al 2019; Zhang et al 2018) and it seems to reduce respiration rates (Yang et el 2022; Zhang et al 2018). Less severe interventions such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018). Negative effects of forest management on soil could be buffered by the addition of forest residues on soil (i.e., slash addition), which can improve the microclimatic conditions, nutrient fertility and biodiversity (ref). We do not know whether the response of soil CO2 effluxes to the microclimatic conditions (i.e. soil temperature and moisture) differs among different forest management strategies. Understanding the sensitivity of respiration to microclimate under different forest management strategies is crucial in a context of climate-smart forestry (CSF) development.

>>>>

Forest disturbances have the potential to affect respiration rates and hence carbon cycling by modifying the microclimate as well as the decomposer communities (ref). Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it due to a reduction of the transpiration of water by trees (Zhang et al 2018). The degree of the response of soil is dependent on the severity of the forest disturbance. Thus, more severe disturbances such as clear-cuts seems to increase soil moisture, enhance heat and frost damage (Kovaks et al 2019; Zhang et al 2018) and reduce respiration rates (Yang et el 2022; Zhang et al 2018). Less severe disturbances such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018). Autotrophic respiration is very sensitive to the decrease in photosynthetic activity due to the loss of trees (Levy-Varon et al. 2012, 2014;Moore et al. 2013; Borkhuu et al. 2015). Instead, increases in heterotrophic respiration due to the increases of saprotrophic activity after the disturabcne

Negative effects of forest disturbances on soil could be buffered by the addition of forest residues on soil (i.e., slash addition), which can improve the microclimatic conditions, nutrient fertility and biodiversity (Pérez-Izquierdo et al. 2025). There is a lack of knowledge about the response of soil CO2 fluxes to the microclimatic conditions (i.e. soil temperature and moisture) under different forest disturbances and if this response can be modulated by effective post-disturbance management strategies. Understanding the sensitivity of respiration to microclimate under different forest disturbances is crucial in a context of climate-change-mediated increased disturbances.
